On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-01-09 manuscript.Rmd
Nico figure added; some writing

@hanson2016 large effect of small gmin on estimates of mesophyll conductance, Ci. This is important but not the focus of this review.

@@

\
The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is interesting because it more closely simulates actual field conditions for plants that are severely droughted, and includes all surfaces of the leaf. Since the seminal review by @kerstiens1996, no attempt has been made to synthesize the literature for values of g~min~, which is timely as models start to incorporate more detailed drought dynamics all the way to the point of mortality by desiccation (Xu, Blackman, Martin-StPaul, Sperry?). Instead, many models rely on inaccurately estimated 'minimum' conductance from cuvette data, by applying regression techniques.

@@

\
(*expand*) We thus have several different definitions of what constitutes a 'minimum' conductance of leaves. In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values to the minimum conductance. We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method only as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].

\
(*expand*) Previews reviews have focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles (riederer, kerstiens). Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between dizzying level of detailed understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models We aim to improve the connection somewhat, and point at obvious gaps in understanding.

@@

\
Measurements of (presumed) minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata (g~cut,seal~) [@kerstiens1996]. The method we here refer to as minimum conductance is that of mass loss of intact detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. Finally, the g~0~ parameter, which is a *modelled* minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

@@

- Measurement error, small fluxes

@@

\
In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated. We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied.

@@

@@

@@

\
Literature values of g~cut~, especially when measured on isolated cuticles (g~cut,isol~) were much lower than any other estimate of conductance (Fig. 1), in line with previous comparisons [@kerstiens1996]. The conductance of stomata-free leaf cuticles represents the absolute minimum conductance that leaves can attain. This result should be viewed with caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. Boyer et al. 1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases (Boyer et al., 1997)

\
Indeed, some more recent studies have shown a very good correspondence between g~cut~ and g~min~ (riederer, schuster). A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.

@@

\
It is important to bear in mind other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, and various interactions with insects [@mueller2008] [@kerstiens1996c;@riederer2008]. The thickness of the cuticle per se may also have to do with mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because  exposure to wind has been shown to dislodge cuticular compounds by abrasion [@rogge1993], thereby increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation could not explain variation in g~cut~. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

\
(*expand*) Water travels more freely across the cuticle than CO2 [@boyer1997; @boyer2015]. This observation was used by @manzoni2011 to recalculate leaf water-use efficiency for droughted plants, but we caution against this because g~cut~ is only one component of g~min~, and CO~2~ will travel through leaky and incompletely closed stomata (see Section XXX). Nonetheless it is an interesting observation that could have major consequences for modelling and interpreting water-use efficiency data during drought conditions. Also @hanson2016 showing the effect of this on calculation of Ci, and calculation of gmeso.

@@

```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
plotfig("output/figure5.png")
```

**Figure 5**. Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves. Left panel, probability density function of all values (after averaging by species, n=200+), with a normal distribution curve (mean=, sd=, on the log scale). Middle panel, g~min~ averaged by phylogenetic order (including only the top ten orders in the database). Bars are 95/% confidence intervals. Numbers on top refer to the number of species, different letters denote significant differences (at alpha = 0.05, adjusted for multiple comparisons), and grey symbols are individual species. Right panel, data for crops only where each point is a genotype. Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species.

@@

```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure6.png")
```

**Figure 6**. Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N). For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).

@@

**Figure 7**. Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C). Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.

>>>>

@@

\
The absolute minimum attainable water loss rate through the cuticle only, which is typically measured on isolated cuticles of the adaxial (non-stomatal) side of the leaf. Other measurements allow for the estimation of water loss through incompletely closed and broken stomata. Of particular interest is the method where leaves are detached from the plant, and their weight loss monitored over time. This method (mass loss of detached leaves, MLD) is of particular interest because it more closely simulates actual field conditions for plants during severe drought, and includes both surfaces of the leaf - not just the stomata-free side which is typically used for cuticular conductance measurements.

@@

\
We thus have several different definitions of what constitutes a 'minimum' conductance of leaves. In this review, we demonstrate that inclusion of a minimum conductance is necessary both for simulation of the response of water-use efficiency to various environmental drivers, as well as a component of the response of plants to severe drought. However, there is no guidance at the moment on how to estimate minimum conductance for use in models, how various definitions compare quantitatively, and whether it is feasible to assign species-specific values of the minimum conductance. We do not expect that g~0~, cuticular conductance (g~cut~) and minimum conductance from mass loss of detached leaves (we refer to minimum conductance measured with this method further as g~min~) to give similar values, in fact the contribution of stomata to each of these conductances will be very different [@brodribb2014].

\
Previews reviews of the minimum conductance have largely focused on the biology of the plant cuticle, and detailed physiology and anatomy of water transport across cuticles [@riederer2008, @schuster2017; @kerstiens1996; @fernandez2017]. Other work has focused on the role of g~0~ in models of plant water use [e.g. @barnard2013; @lombardozzi2017], without referring to the literature on cuticular conductance. Thus there is a disconnect between the dizzying level of understanding of the plant cuticle [@riederer2008], and the very simple assumptions made when using leaf conductance models in global vegetation models. We aim to improve this connection by synthesizing the state of knowledge, thereby pointing out gaps in understanding, and suggest areas for further research.

@@

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories (see Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cut~) are typically made on isolated non-stomatal cuticles [@riederer2001], via gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata (g~cut,seal~) [@kerstiens1996]. The method we here refer to as minimum conductance is that of mass loss of intact detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. After some time, a constant minimum transpiration rate is attained (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), from which minimum conductance (g~min~) can be inferred using measured vapour pressure deficit [@sinclair1986]. This method most closely resembles conditions that plants may experience during a dry-down or periods of extreme stress. Finally, the g~0~ parameter, which is a model-based minimum conductance reached when photosynthesis rate is zero, is typically estimated by applying regression to Eq. 1 (of which various forms exist).

@@

@@

\
In the following we demonstrate that the fitting process is problematic for three reasons: 1) the estimates of g~0~ and g~1~ are highly correlated, 2) the precision for g~0~ is generally much lower than for g~1~, and 3) for data that have a worse fit overall, the g~0~ estimates are elevated. We thus conclude that g~0~ should really not be estimated from regression, although this is the most common method applied. In addition, it is difficult to accurately measure low fluxes with a portable gas exchange system, and great care has to be taken to arrive at reasonable estimates of g~0~ that way.

@@

\
In this review we focus on the cuticle as a barrier to water loss, but point out that previous work has shown that the cuticle is much less permeable to CO~2~ compared to H~2~O [@boyer1997; @boyer2015]. This finding, if indeed generally true, has a large effect on the calculation of various gas exchange parameters [@hanson2006]. @manzoni2011 recalculated leaf water-use efficiency for droughted plants assuming CO~2~ is blocked by the cuticle, but the uncertainty of this approach is leaky and incompletely closed stomata (see Section XXX). Because g~0~ includes not just the cuticle, but also incompletely closed stomata, and the finding that g~0~ is much larger than estimates of g~min~ or g~cut~ (Fig. 1), we ignore this effect in the model simulations that follow.

@@

```{r fig.height=6, echo=FALSE, eval=includefigs}
plotfig("output/figure5.png")
```

**Figure 5**

@@

\
Literature values of g~cut~, especially when measured on isolated cuticles (g~cut~) were much lower than any other estimate of conductance (Fig. 1), in line with the methods comparison of @kerstiens1996. More recently, @schuster2017 argued, based on a data for 8 species, that g~min~ from MLD is comparable to g~cut~. However in the same study, a literature compilation of hundreds of values showed that g~cut~ was on 10-fold lower compared to g~min~, though this finding was not reported by @schuster2017 [see also @fernandez2017], and instead g~min~ and g~cut~ were pooled in their analysis. Nonetheless, the comparison between g~cut~ and g~min~ should be viewed with some caution as @kerstiens1996 argued that measurements of g~cut~ maybe too low due to low water content of the cuticle after detachment from the leaf, and storage in often dry-air conditions. @boyer1997 suggested that this decline of the cuticular permeability could be due to stretching of the wax layer at full saturation and tightening of the wax structure as the turgor releases.

\
A review of cuticular transport mechanisms, biochemical composition, and formation of cuticular waxes is well outside the scope of this review, as these processes have been well described elsewhere [@kerstiens1996; @kerstiens2006; @shepherd2006; @schreiber1996; @schreiber2001; @schuster2016b]. However, a few key points should be reiterated as they are relevant to the current discussion, in particular when we aim to interpret variation in literature values of g~min~.

@@

\
It is important to bear in mind the many other functions of the cuticle, including attenuation of radiation, as a barrier for fungal pathogens, various interactions with insects including signalling and herbivory resistance [@mueller2008; @kerstiens1996c;@riederer2008]. The thickness of the cuticle may also confer mechanical strength, as @onoda2012 reported that thicker cuticles are more resistant to tearing (the force to tear and tensile strength both approx proportional to cuticle thickness). Thick cuticles may also be useful because exposure to wind has been shown to dislodge cuticular compounds by abrasion [@rogge1993], thereby increasing the conductance of the cuticle [@hadley1983;@grace1974;@vangardingen1991]. Similarly, exposure to simulated rain damages the cuticle and increases its conductance [@baker1986], most dramatically demonstrated in an ice storm, leading to much higher g~min~ [@boyce2003]. Such exposure to the elements was also held responsible for thicker cuticles in coastal, wind-exposed sites compared to inland sites [@blackman2005].

\
Despite considerable work on the topic, no clear relationship between chemical composition or structure of the cuticle and its conductance has emerged. Recently, @schuster2016a presented a comprehensive study of cuticle chemical composition, but this variation explains only some of the wide range of g~cut~. Similarly, @hauke1998 found no relationship between the gradual decrease in g~cut~ in *Hedera helix* during leaf maturation and cuticle wax amount, mean chain lengths, or cuticle weight.

@@

```{r fig.height=5, fig.width=5, echo=FALSE, eval=includefigs}
plotfig("output/figure6.png")
```

**Figure 6**. Analysis of a new literature compilation of g~min~ estimates, as measured with mass loss of detached leaves. Left panel, probability density function of all values (after averaging by species, n=200+), with a normal distribution curve (mean=, sd=, on the log scale). Middle panel, g~min~ averaged by phylogenetic order (including only the top ten orders in the database). Bars are 95/% confidence intervals. Numbers on top refer to the number of species, different letters denote significant differences (at alpha = 0.05, adjusted for multiple comparisons), and grey symbols are individual species. Right panel, data for crops only where each point is a genotype. Bars denote the range, illustrating the large variation in g~min~ between genotypes for a particular crop species.

@@

**Figure 7**. Minimum conductance measured from leaf dry-down curves on a variety *Hakea* species, a genus native to Australia. Plants were grown in containers in a grow house, and supplied with ample water or underwent long-term mild drought stress. Bars are labelled by leaf form : broadleaf (B) or needle-like (N). For droughted plants, g~min~ was higher for species with needle-like leaves (Wilcoxon test, p `r pval(wilcox.test(gmin.d ~ leaf.form, data=lopw)$p.value)`)), but not for well-watered plants (p `r pval(wilcox.test(gmin.w ~ leaf.form, data=lopw)$p.value)`).

@@

```{r fig.height=4,echo=FALSE, eval=includefigs}
plotfig("output/figure8.png")
```

**Figure 8**. Minimum conductance measured from leaf dry-down curves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in a constant elevated temperature treatment (+3 degrees C). Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There is no effect of measurement T on g~min~, but leaves of trees grown in elevated temperature showed consistently lower g~min~.
